### 2025-04-17 docs/Homeownership_BankOfMomAndDad.tex
Added preliminary results on effect of transfers

@@

\begin{table}[htbp]
	\begin{adjustbox}{center}
		\small
		\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
		\begin{threeparttable}
			\singlespacing
			\caption{Descriptive Statistics (Means), Households Aged 25-45}\label{tab:descrstats}
			\begin{tabular}{l  r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r}
				\toprule
				& \multicolumn{2}{c}{All}  &  \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver 		& \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\  \midrule
				\input ../tabfig/descr/descr_trsnf_disc_mean.tex
				\midrule
				Age                    & \multicolumn{2}{c}{25-31} & \multicolumn{2}{c}{32-38}  & \multicolumn{2}{c}{39-44}   \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				 \midrule
				\input ../tabfig/descr/descr_trsnf_age_mean.tex
				\midrule
				Wealth Tertile                    & \multicolumn{2}{c}{Tertile 1}  & \multicolumn{2}{c}{Tertile 2}  & \multicolumn{2}{c}{Tertile 3}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				\midrule
				\input ../tabfig/descr/descr_trsnf_wlt_mean.tex
		\bottomrule
			\end{tabular}
			{\footnotesize
				\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD. }
		\end{threeparttable}

@@

I now utilize microdata from the PSID to examine the correlation between parental financial resources and housing outcomes.

@@

@@

>>>>

\textcolor{red}{In all other waves, there is a general question of gifts and inhertitances from people outside the hosuehold. We do not know the source of these transfers (e.g., they could be from friends or an inheritance from grandparents) and hence these do not know the extent to which these captures parental assistance. Yet, they, to some extent, capture the impact of transfer receipt on household decisions, and hence I use these variables to capture the role of transfers on housing outcomes, following \cite{Lee2018}.}

@@

\begin{table}[htbp]
	\begin{adjustbox}{center}
		\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
		\begin{threeparttable}
			\singlespacing
			\caption{Descriptive Statistics (Means), Households Aged 25-44}\label{tab:descrstats}
			\begin{tabular}{l  r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r r@{\hspace{1.\tabcolsep}}r}
				\toprule
				& \multicolumn{2}{c}{All}  &  \multicolumn{2}{c}{Renter}   & \multicolumn{2}{c}{Owner}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver 		& \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\  \midrule
				\input ../tabfig/descr/descr_trsnf_disc_mean.tex
				\midrule
				Age                    & \multicolumn{2}{c}{25-31} & \multicolumn{2}{c}{32-38}  & \multicolumn{2}{c}{39-44}   \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				 \midrule
				\input ../tabfig/descr/descr_trsnf_age_mean.tex
				\midrule
				Wealth Tertile                    & \multicolumn{2}{c}{Tertile 1}  & \multicolumn{2}{c}{Tertile 2}  & \multicolumn{2}{c}{Tertile 3}    \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
				Receiver       & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes} & \multicolumn{1}{c}{No} & \multicolumn{1}{c}{Yes}    \\ 
				\midrule
				\input ../tabfig/descr/descr_trsnf_wlt_mean.tex
		\bottomrule
			\end{tabular}
			{\footnotesize
				\textit{Notes:} Data from the PSID Transfer, Individual, and Family modules. Weighted using family weights. Transfer, wealth, and income measured in 1000s of 2016 USD. }
		\end{threeparttable}

@@

I now utilize microdata from the PSID to examine the empirical association between transfer receipt and parental financial resources on housing outcomes.

@@

\subsubsection{Transfers and the Rent-to-Own Transition}
\textcolor{red}{!! [[rewrite this]] !!}
As discussed in the introduction, several papers look at the role of parental transfers and resources for entering into the housing market \cite[see e.g.,][]{wold2024housing,Lee2019,Blickle2019}. The goal of this section is not to `re-invent' the wheel, but to simply show that these relationships look like in my data. I run regressions on renting households aged 25-44 who have not been observed as homeowners before. In the first set of regressions, the sample are all households aged 25-44 in 2011 who are renting with non-missing parental transfer data from the 2013 Transfer supplement. The outcome variable is whether the household is an owner in the next period, and the set of control variables are (log) parental cash-on-hand, (log) household wealth, (log) household income, household marital states, (log) household size, education, age, parental age, and fixed effects for race, year, and state. The results of the transfer coefficients are reported in Table \ref{tab:newowners_short} while the full regression output is in Appendix Table \ref{tab:newowners}

Column 1 reports the effect of recieving a parental transfers (above \$500) (in year 2012) on the probability of becoming a homeowner in 2013, which increases by 1.6 pp. Next, I also include the inheritance/transfer dummy, (hence includes households who received transfers/inheritances above \$10,000 in 2011, 2012, or 2013). Column 2 also includes households who report having recieved transfers or inheritances above \$10,000 while column (3) only denotes those who have received transfers above \$10k (whether transfers from parents or any inheritance). Renters who receive transfers are much more likely to transition to homeowners, even after controlling for other observable characteristics. If we only include large transfers (above \$10,000) transfer receipt is associated with a 7.7 percentage point increase in the probability of transitioning into homeownership, more than half of the transition rate. The coefficents are not statistically significant, perhaps not surprising given the small sample size, the few households who receive transfers and the few households who become homeowners.

To increase statistical power I run the same regression using the same specification, but over the entire sample period (1999-2021) using OLS and household fixed effects (FE). Using fixed effects also allows us to control for household fixed effects such as genetics, risk aversion, etc. The downside is that we only observe larger transfers (above 10k) and that we do not know the source of the transfer. If we focus on the FE estimator, we see that transfer receipt is associated with an 8.3 percentage point increase (statistically signifacant at the 10\% level) in the probability of becoming a homeowner, relative to a benchmark probability of 18.1\%.

It is perhaps not so surprising that the effect of transfer receipt is so big. First, beyond the 2013 transfer supplement (which has the smallest effect, 1.6 pp) I only capture transfers above \$10,000. Standard models, including the quantiative model in this paper, predicts that renters follow a threshold rule: for any renter that chooses to rent, there exists a wealth level at which they prefer to own (Indeed, the regression results also show that a 10\% increase in net worth increases the probability of becoming a homeowner by 4 percentage points.). Hence, large transfers can shift households above the trehsold if they were `close'. Second, these transfers are not exogenous: indeed, it is likely that some of these transfers given exactly because the household is buying.

\begin{table}
	\center
	\begin{threeparttable}
		\caption{The Transition to Ownership}
		\label{tab:newowners_short}
		\small 
				\input{../tabfig/regr/newowners_short.tex}

		{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses. \textcolor{red}{update footnote} \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}		
	\end{threeparttable}
\end{table}

To illustrate the size issue and data limitations, and the need for a model, I plot the size of transfers to renters, by whether they become homeowners in $t+2$ in Figure \ref{fig:PSID_transfer_to_renters}. The left planel only includes transfers from parents to children (from the transfer supplement) in 2013 while the right panel plots the data from the transfer/inheritance above 10k figure. There are several things to note. First, we note that, among receivers, households who transition to ownership receive larger transfers on average, in both figures. Second, the transfers in the transfer supplement is much smaller than than those in the gift/inheritance question. Indeed, the median parental transfer is just about \$1k, the mean about \$3k, and the 95th percentile is at \$10k. Third, In all measures, no matter the measure, we see that receiving renters who become owners receive much larger transfers. 
\begin{figure}
	\caption{Transfer Sizes to Renters by Transition to Ownership}\label{fig:PSID_transfer_to_renters}
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_transfers.pdf}% 
	\includegraphics[width=0.5\textwidth]{../tabfig/descr/PSID_transfers_renters_inheritance.pdf}%

	 {\begin{footnotesize} \textit{Notes:} The left panel uses the transfers, gifts, and inheritance variable which is available in all waves of the PSID while the right panel uses the parental transfer data only available in the 2013 transfer supplement. See main text for a discussion. Vertical dotted and dashed lines denote the median and means, respectively. The figures show the kernel density estimates by group, capped at the the 99th percentile. \end{footnotesize}}
\end{figure}

@@

\begin{table}
	\center
	\begin{threeparttable}
		\caption{The Transition to Ownership}
		\label{tab:newowners}
		\small 
				\input{../tabfig/regr/newowners.tex}

		{\begin{footnotesize}\begin{flushleft}
		\textit{Notes:} Standard errors in parentheses. \textcolor{red}{update footnote} \textsuperscript{+}$p<0.10$, \textsuperscript{*}$p<0.05$, \textsuperscript{**}$p<0.01$, \textsuperscript{***}$p<0.001$.
		\end{flushleft}\end{footnotesize}}		
	\end{threeparttable}
\end{table}
